\addchaptertocentry{Acknowledgement}

 

I dedicate this thesis to my parents who are the reason I was able to get this far. My father’s generous spirit and my mother’s indomitable attitude have allowed me to face life’s challenges with dignity. I owe them far more than I can ever repay.

Though only my name appears on the cover of this thesis, a great many people have contributed to its production. My deepest gratitude is to my guide, Dr. Solomon FD Paul. I have been amazingly fortunate to have a guide who gave me the freedom to explore on my own and taught me how to express ideas. His patience and support helped me overcome many crisis situations and finish this thesis. I will forever remain thankful for allowing me to be me.

I would not like to think of where I would be today if it weren’t for my three pillars of support. Dr. P. Venkatachalam, has been always there to listen and give advice. My heartfelt gratefulness goes out to him for time spent on helping me understand research, reviewing my manuscripts and enriching my ideas. His work and research ethic will forever be a benchmark for me to attain. Dr. V. Vettriselvi is an irreplaceable friend and mentor. I am deeply grateful to her for the long discussions that helped me sort out the details of my work. She has taught me the importance of patience and remaining calm in the face of any adversity. Dr. N. Venkateswaran has been an unlimited source of humor and resources. His unending optimism has seen me through many a dark day.

I owe a debt of gratitude to my research advisory committee member Dr. Sridevi Hegde for taking a chance on me when I was a fresher looking for a job. She continues to be my confidante and an inspiration. Dr. Kalpana Gowrishankar has been a tremendous source of advice often guiding the direction I took my thesis with her insights. 

I thank the management and authorities at Sri Ramachandra University for providing facilities and opportunities for members of the faculty to pursue a higher degree. I am particularly grateful to Dr SP Thyagarajan, Professor of Eminence \& Dean, SRU, for granting me permission to utilize the Central Research Facility. 

I will forever be thankful to Dr. Jebaraj (SRMC), Dr. Suresh and Dr. Rajan (Madras Medical Mission) who helped me with the sample collection and their clinical inputs.  I am truly obliged to all the volunteers who graciously accepted to be a part of this research. Despite the emotional trauma most of these families were going through, they didn’t hesitate to be a part of this research which is a testimony to their generous hearts. 

My sincere gratitude to Dr. P. Venkatesan who always accommodated me for my statistical analysis without hesitation, and Dr. Vikram Jayanth who always reminded me of the importance of prioritizing and time management. I value the endless pep talks and advice given by Dr. Santhi S and Dr. Mangathayaru K, urging me to complete my degree. Dr. Santhi’s timely help with the PCR is greatly appreciated. I thank all my colleagues, Dr M Ravi, Dr. V Ganesh , Dr J Vijayalakshmi , Dr Harpreet Kaur, Ms Andrea Mary, Ms R Selvi, Ms Deepa Parvathy Koushik , Ms Nandhini B, and staff, Ms Sujikumari, Ms Lalitha and Ms Saraladevi, all of whom have been a part of my journey in their own unique way.

Many friends have helped me stay sane through these difficult years. Their support and care helped me overcome setbacks and stay focused. I greatly value their friendship and I deeply appreciate their belief in me. Topping that list is soon to be Dr.s’ Ganga and Shruthi. Thank you for having me as your pillion rider all these years and making it possible for me to stay late and finish my work. Trying to teach me work-life balance was an uphill battle for Ganga, one that she has almost won. So if I can smell the roses today it’s all because of her. If the thesis looks good, I owe it all to Shruthi who spent sleepless nights doing the Latex typesetting and never desisted till it was perfect. A heartfelt thank you to my students who remained my friends; Ms. Kamna R who helped me with the images, Ms. Shruti Daga and Ms. Sarada for their assistance in sample collection and Ms. Raheema Beevi who insisted on helping me out with the DNA isolation. Ms. Priyanka and Ms. Mohanapriya have been my lab partners in an inexplicable way. I guess no one else would understand the ``50\% success'' more than these two. I thank Dr. K. Vijaylakshmi and Mr. Saurabh, both of whom have always been a source of confidence, proving that distance does not matter when it comes to friendship. Last but most certainly not the least, I am blessed to have my surrogate family, the Daniels, with me all these years. Their quiet support and encouragement has meant more to me than I can express. 

I owe my gratitude to all those people who have made this thesis possible and because of whom my research experience has been one that I will value forever.


\vspace{4em}
\begin{center}
\begin{minipage}{0.6\textwidth}

 ``Knowledge is in the end based on acknowledgement.'' \\
― Ludwig Wittgenstein

\end{minipage} 
\end{center} 
