% Appendix Template

\chapter{Publications \& Presentations} % Main appendix title

\label{Appendix1} % Change X to a consecutive letter; for referencing this appendix elsewhere, use \ref{AppendixX}

\textbf{Shruthi Mohan}, Teena Koshy, Perumal Vekatachalam, Sheela Nampoothiri, Dhanya Yesodharan, Kalpana Gowrishankar, Jeevan Kumar, Latha Ravichandran, Santhosh Joseph, Anupama Chandrasekaran, Solomon FD Paul. Subtelomeric rearrangements in Indian children with idiopathic intellectual disability/developmental delay: frequency estimation and clinical correlation using FISH.
In press: Indian Journal of Medical Research 

\textbf{Shruthi Mohan}, Vettriselvi Venkatesan, Teena Koshy, Solomon FD Paul, Perumal Venkatachalam. Genomic imbalances in subjects with idiopathic intellectual disability detected by multiplex ligation-dependent probe amplification.
In press: Journal of Genetics

\textbf{Shruthi Mohan}, Sheela Nampoothiri, Dhanya Yesodharan, Vettriselvi Venkatesan, Teena Koshy, Solomon FD Paul, Perumal Venkatachalam. Reciprocal microduplication of the Williams-Beuren syndrome (7q11.23) region. 
Under review: Lab Medicine

Teena Koshy, Vettriselvi Venkatesan, Kalpana Gowrishankar, Venkatachalam Perumal, \textbf{Shruthi Mohan}, Solomon FD Paul. Mutation analysis of Tbx1 in children with conotruncal heart anomalies from a referral hospital in South India. 
Under Review: Indian Journal of Pediatrics

Shruthi Mohan, Teena Koshy, Solomon FD Paul, Latha Ravichandran, Venkatachalam P. Detection of subtelomeric rearrangements in idiopathic mental retardation by fluorescence in situ hybridization. 5th Annual Conference on Genetic and Molecular Diagnosis in Modern Medicine, SRM, Chennai; 2012.

Shruthi Mohan, Teena Koshy, Venkatachalam P, Solomon FD Paul, Latha Ravichandran, Kalpana Gowrishankar, Sheela Nampoothiri, Raheema Beevi, Suji Kumari. Subtelomeric rearrangements in children with idiopathic mental retardation. Next Revolution in Genetics and Genomics, Sir Gangaram Hospital, New Delhi; 2013.

